---
title: "Strictly Hypothesis Testing"
author: "Doug Ashton"
date: "25/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

I'll get straight to it. I love [Strictly](https://www.bbc.co.uk/strictly). You may know it by "Dancing with the Stars", "Bailando por un Sueño", or "Danse avec les stars". I resisted for a decade, but for the last two years I've been fully seduced by the big glitter ball and haven't missed an episode.

Now as a fully qualified arm-chair critic, I think I've noticed that to get a good score on Strictly you just need to do the Charleston. It hides a multitude of sins. You sort of jump about a bit, do something called "swivel", and fall over at the end. Then you get a minimum 35. Job done.

My family disagree, so I've set about to test my hypothesis. The great news is that this is really easy to do with R. In fact, this whole post uses no more R than you'd find in an [introduction to R course](https://www.mango-solutions.com/additional-solutions/r-training). My family can't wait to find out the results. Wherever they went.

For the rest of this post you can relive with me how I tested out my idea and finally put this to rest. Base R has everything we need but we'll get there quicker with the [tidyverse](https://www.tidyverse.org/).

```{r libraries}
library(tidyverse) # I've set message=FALSE everywhere
```

## The Data

We're lucky that someone at [www.ultimatestrictly.com](https://www.ultimatestrictly.com/) has been keeping track of every dance that's ever happened on strictly. And we're even luckier that they have made it available as a CSV download. It only goes up to 2016 but that's still 14 series so it should be enough.

With the URL all we need are a couple of tweaks to the default `read_csv` to account properly for an unusual NA string and some sparse columns that guess the data type incorrectly.

```{r data, cache=TRUE}
url <- "https://www.ultimatestrictly.com/s/SCD-Results-S14.csv"
raw_results <- read_csv(file = url, na = c("-"), guess_max = 10000)
```

Let's start with a little light cleaning. The formatting on the dances is a bit inconsistent. Most of it can be fixed by forcing the dance names to be a consistent case. I like the look of title case so we'll use stringr's `str_to_title`.

```{r cleaning}
dances <- raw_results %>%
  mutate(Dance = str_to_title(Dance))

head(dances)
```

So you see we get every dance, and individual judges scores as well as the total score. My next step would usually be to [gather](https://tidyr.tidyverse.org/reference/gather.html) this into a tidy long data frame, maybe [separate](https://tidyr.tidyverse.org/reference/separate.html) the couples into celebrity and professional. But in this instance we only want the Dance and the Total, so we'll leave it like it is.

## Top Dances

I'd like to know which dances get the best `Total` score. We can do this with a `group_by`-`summarise` combination.

```{r topdances1}
dances %>%
  group_by(Dance) %>%
  summarise(MeanScore = mean(Total, na.rm = TRUE),
            Count = n()) %>%
  head(3)
```

I'm removing anything with less than 10 scores because these tend to be hybrid dances and one-offs.

```{r nosparse, eval=FALSE}
  filter(Count >= 10) %>%
```

The [Show Dance](http://www.ultimatestrictly.com/show-dance/) only appears in the final episode and is multi-style. Out it goes. 

```{r noshow, eval=FALSE}
 filter(Dance != "Show Dance") %>%
```

I don't want any "hops" in there. The hops are those ones where everyone's on the floor at the same time and you have no idea what's going on but they all seem to be enjoying themselves. There's only a Lindy Hop in this data but I don't want any hops. Ever. No Hops!

```{r nohop, eval=FALSE}
  filter(!str_detect(Dance, " Hop")) %>%
```

Putting it all together we get our top dances.

```{r topdances}
top_dances <- dances %>%
  group_by(Dance) %>%
  summarise(MeanScore = mean(Total, na.rm = TRUE),
            Count = n()) %>%
  filter(Count >= 10)  %>%
  filter(Dance != "Show Dance") %>%
  filter(!str_detect(Dance, " Hop")) %>%
  arrange(desc(MeanScore))

knitr::kable(top_dances)
```

And there we have it. I'm sort of right. The Charleston is top 3. And who doesn't like an Argentine Tango? But there's something about that count... What if they tend to only do the Charleston in later weeks when they're a bit better? 

We know that the scores go up by week. In fact it's pretty linear.

```{r weekplot, warning=FALSE}
dances %>%
  ggplot(aes(x = Week, y = Total, group = Week)) + 
  geom_boxplot()
```

Let's do the same again and add average week.

```{r topdancesweek}
top_dances_week <- dances %>%
  group_by(Dance) %>%
  summarise(MeanScore = mean(Total, na.rm = TRUE),
            MeanWeek = mean(Week, na.rm = TRUE),
            Count = n()) %>%
  filter(Count >= 10)  %>%
  filter(Dance != "Show Dance") %>%
  filter(!str_detect(Dance, " Hop")) %>%
  arrange(desc(MeanScore))

knitr::kable(head(top_dances_week, 5))
```

Rats. My hypothesis is in trouble. My children have long since gone to bed so let's take it up a notch and build a statistical model that can account for `Week` and `Dance` at the same time.

## Linear Model

First, we'll take the individual dance level data and reuse the filtering for the `top_dances`.

```{r modeldata}
main_dances <- dances %>%
  filter(Dance %in% top_dances$Dance)
```

We want a simple linear model that accounts for the increasing score by week and the effect we're interested in -- the dance.

```{r lm}
fit <- lm(Total ~ Week + Dance, data = main_dances)
summary(fit)
```

After removing the strong effect of week, there's not much left of significance. It can be tempting to cherry pick individual levels but the only one with much signal is the Cha Cha Cha. Ultimate Strictly describes it as ["A cheeky, fun dance, but rarely a show stopper"](https://www.ultimatestrictly.com/cha-cha-cha/).

## Conclusion

Unfortunately it looks like the evidence doesn't support my hypothesis that the Charleston is more favourably marked. Maybe all that swivelling is harder than it looks. Faye and Giovanni did undeniably do an excellent job with their, perfect scoring, [Sound of Music Charleston](https://youtu.be/55z6IrL8k0Q). What we did find is that the Cha Cha Cha might score lower. Although that's most likely because it's seen as an easier dance for. Well. People like Quentin.

<iframe width="560" height="315" src="https://www.youtube.com/embed/Da_zxblfGoI?start=126" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

I'm sorry Quentin. I would fare no better.

Well seeing as I'm wrong, and everyone else is asleep. Let's just pretend this never happened.


[Code for this blog post on GitHub](https://github.com/mangothecat/blog_strictly).
